### 2016-01-14 modules/1_structure.tex
Restructered the whole project. Now built on multi page setup. Added an image and some refs.

\documentclass[../main.tex]{subfiles}
\begin{document}

\chapter{Überblick der Arbeit / Struktur der Arbeit}
\label{structure}

\end{document}