### 2016-01-14 modules/abstract.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% Abstract
\begin{abstract}
\label{abstract}
  \noindent\emph{English version:}\newline\newline
  \noindent
  ....\\
  ..\\

  \vspace{1cm}
  \noindent\emph{Deutsche Version:}\newline\newline
  \noindent
  ....\\
  ...\\
  .\\
  ..\\
\end{abstract}

\end{document}