### 2016-03-15 modules/hyphenation.tex
Tried to do longtables multipage feature right but failed. Reduced table from 2_basics so it fits on one page now. Added hyphenation plain text file. Added shortcuts cvalid and cbroken. Minor changes.

% Reference: https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_Silbentrennung

\hyphenation{Zu-wanderung}