### 2016-03-16 modules/hyphenation.tex

% Reference: https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_Silbentrennung

\hyphenation{}