### 2016-03-19 modules/hyphenation.tex

% Reference: https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_Silbentrennung

\hyphenation{}

% Integrationsmöglichkeiten
% Optimierungsmöglichkeiten
% übergeordneten