### 2016-03-27 modules/acronyms.tex

% ACRONYMS %

\newacronym{OS}{OS}{Operating System}
\newacronym{IT}{IT}{Informationstechnik}
\newacronym{BSD}{BSD}{Berkeley Software Distribution}
\newacronym{CPU}{CPU}{Central Processing Unit}
\newacronym{VM}{VM}{Virtual Machine}
\newacronym{UI}{UI}{User Interface}
\newacronym{API}{API}{Application Programming Interface}
\newacronym{REST}{REST}{Representational State Transfer}
\newacronym{OSI}{OSI}{Open Systems Interconnection (Modell)}
\newacronym{COW}{COW}{Copy-On-Write}
\newacronym{LSM}{LSM}{Linux Security Modules}
\newacronym{MAC-SEC}{MAC}{Mandatory Access Control} %  (Zugriffskontrolle)
% \newacronym{MAC-NET}{MAC}{Media Access Control (Netzwerkkommunikation)}
\newacronym{RBAC}{RBAC}{Role-Based Access Control}
\newacronym{LXC}{LXC}{Linux Containers}
\newacronym{SELinux}{SELinux}{Security Encanced Linux}
\newacronym{AppArmor}{AppArmor}{Application Armor}
\newacronym{JSON}{JSON}{JavaScript Object Notation}
\newacronym{MITM}{MITM}{Man-In-The-Middle}
\newacronym{CISSP}{CISSP}{Certified Information Systems Security Professional}
\newacronym{OCF}{OCF}{Open Container Format}
\newacronym{OCP}{OCP}{Open Container Project}
\newacronym{runC}{runC}{runContainer}
\newacronym{IP}{IP}{Internet Protocol}
\newacronym{HTTP}{HTTP}{Hypertext Transfer Protocol}
\newacronym{HTTPS}{HTTPS}{Hypertext Transfer Protocol Secure}
\newacronym{TLS}{TLS}{Transport Layer Security}
\newacronym{NSA}{NSA}{National Security Agency}
\newacronym{PID}{PID}{Process ID, Process Identifier}
\newacronym{UTS}{UTS}{UNIX Time Sharing}
\newacronym{ARP}{ARP}{Address Resolution Protocol}
\newacronym{MCS}{MCS}{Multi-Categegory Security}
\newacronym{MLS}{MLS}{Multi-Level Security}
\newacronym{rlimits}{rlimits}{Resource Limits}
\newacronym{cgroups}{cgroups}{Control Groups}
\newacronym{I/O}{I/O}{Input and Output}
\newacronym{HDD}{HDD}{Hard Disk Drive}
\newacronym{SSD}{SSD}{Solid State Drive}
\newacronym{USB}{USB}{Universal Serial Bus}
\newacronym{ACL}{ACL}{Access Control List}
\newacronym{DAC}{DAC}{Discretionary Access Control}
\newacronym{CA}{CA}{Certificate Authority}
\newacronym{CLI}{CLI}{Command-Line Interface}
\newacronym{GUI}{GUI}{Graphical User Interface}
\newacronym{TUF}{TUF}{The Update Framework}
\newacronym{CVE}{CVE}{Common Vulnerabilities and Exposures}
\newacronym{TOML}{TOML}{Tom's Obvious, Minimal Language}
\newacronym{LTS}{LTS}{Long Term Support}
\newacronym{RHEL}{RHEL}{Red Hat Enterprise Linux}

% This syntax is supposed to do cross referencing, but doesn't work with nonumberlist.
%\newacronym[see={[Glossar:]{ContinuousIntegration}}]{CI}{CI}{Continuous Integration}
% Thus manual cross references...
\newacronym{CI}{CI}{Continuous Integration, \textit{s. Glossar:} \gls{ContinuousIntegration}}
\newacronym{CD}{CD}{Continuous Delivery, \textit{s. Glossar:} \gls{ContinuousDelivery}}
\newacronym{IPC}{IPC}{Inter Process Communication, \textit{s. Glossar:} \mbox{\gls{InterProcessCommunication}}} % Visual fix with mbox{} to prevent new line for last syllable.
\newacronym{DoS}{DoS}{Denial of Service, \textit{s. Glossar:} \gls{DenialOfService}}
\newacronym{chroot}{chroot}{change root, \textit{s. Glossar:} \gls{chrootGlossary}}